---
title: "Virtual Lab Session Day 1"
author: "Srijit Mukherjee"
date: "23 June 2020"
output:
  beamer_presentation: default
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, size = "tiny")
knitr::opts_chunk$set(fig.align = "center", fig.height = 3, fig.width = 3)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

## Types of Data Variables in R


- Quantitative

    - Continuous
    - Discrete
      
- Qualitative

    - Nominal
    - Ordinal

## Datasets in R

If you type the following command, you will get to know all the data sets in R.

```{r, echo = TRUE, comment = ''}
data()
```

## Handling Data in R

Let's take the mtcars data set in R. Let's see the top of the data matrix.

```{r, echo = TRUE, comment = ''}
data = mtcars
head(data, n = 3)
```

## Handling Data in R

Let's see the bottom of the data matrix.

```{r, echo = TRUE, comment = ''}
tail(data, n = 3)
```

## Handling Data in R

Let's see a concise summary of the data matrix.

```{r, echo = TRUE, comment = ''}
str(data)
```

## Handling Data in R

```{r, echo = TRUE, comment = ''}
class(data)
class(data$mpg)
dim(data)
names(data)
```

## Continuous Data

Exploratory Data Analysis

```{r, echo = TRUE, comment = ''}
mean(data)
mean(data$mpg)
var(data$mpg)
sd(data$mpg)
```


## Continuous Data

Exploratory Data Analysis

```{r, echo = TRUE, comment = ''}
min(data$mpg)
max(data$mpg)
median(data$mpg)
```

## Continuous Data

Exploratory Data Analysis

```{r, echo = TRUE, comment = ''}
quantile(data$mpg)
range(data$mpg)
summary(data$mpg)
```

## Continuous Data

Exploratory Data Analysis

```{r, echo = TRUE, comment = ''}
summary(data)
```

##Continuous Data

Exploratory Data Analysis

```{r, echo = TRUE, comment = ''}
hist(data$mpg, xlab = "MPG", probability = TRUE, , main = "Histogram")
```


##Continuous Data

Exploratory Data Analysis

```{r, echo = TRUE, comment = ''}
plot(data$mpg, data$wt, xlab = "MPG", ylab = "Weight") 
```

##Continuous Data

Exploratory Data Analysis

```{r, echo = TRUE, comment = ''}
plot(data$mpg, data$hp, xlab = "MPG", ylab = "Horsepower")
```

##Continuous Data

Exploratory Data Analysis

```{r, echo = TRUE, comment = ''}
plot(data$hp, data$wt, xlab = "Horsepower", ylab = "Weight")
```

## The sapply() function

This function helps to apply a function to a subset of the data frame. 

Let's say, you want to compute the mean of the data frame.

```{r, echo = TRUE, comment = ''}
sapply(data, mean)
```

## The Descriptive Statistics package and function

```{r, echo = TRUE, comment = '', eval=TRUE}
#install.packages("pastecs")
library("pastecs")
```

## The Descriptive Statistics package and function

```{r, echo = TRUE, comment = ''}
round(stat.desc(data), 2)
```

## The Box Plot

